---
title: "bioko_macro_workflow"
author: "Daniel T Citron"
date: "1/22/2020"
output: html_document
---


```{r Load Libraries}
library(data.table)
library(ggplot2)
library(Matrix)
library(MASS)

library(here, lib.loc = "libraries/")
library(macro.pfsi, lib.loc = here("libraries/macro.pfsi"))
```

```{r Load data}
data.path <- "/ihme/malaria_modeling/data/BIMEP_2019_Data/data/"
# population data
pop.data <- fread(paste0(data.path, "clean/aggregated_2015_2018_travel_data.csv"))
# travel frequency
trip.freq.data <- fread(paste0(data.path, "clean/trip_frequency_model_estimates.csv"))

# travel destination selection model
trip.dest.data <- fread(paste0(data.path, "clean/negative_binomial_predictions_by_destination_region.csv")) 

# travel duration
trip.duration.eg <- 1/0.04713604 # rate of return from mainland eg to bioko
trip.duration.bi <- 1/0.09670302 # rate of return from trips on bioko

# care seeking behavior
FeverPf = 0.1116336
TreatPf = 0.602

# PfPR data, focusing on the 2018 data first
pfpr.data <- fread(paste0(data.path, "clean/pfpr_draws.csv"))
pfpr.data <- merge(pfpr.data, pop.data[year == 2018, .(areaId)], by = "areaId", all = FALSE)
```

Construct the vector of PfPR inputs - note that we have added one more patch location to this list - this is the mainland.
Overall, there are 242 patches including 241 on island and 1 more off island
```{r PfPR input vector}
# Here's the vector of PfPR that we're calibrating to:
pfpr.input <- c(pfpr.data$draw.mean, 0.43)
print(c(summary(pfpr.input), length(pfpr.input)))
```

Dynamical model parameters
```{r Dynamical model parameters}
a = 0.3*0.9
b = 0.55
c = 0.15
r = 1./200 # rate at which people become cured
eta = 1./30 # rate at which prophylaxis wears off
p = 0.9 # fraction of surviving mosquitoes
g = 1 - p # fraction of dying mosquitoes
peip = p^11 # fraction of mosquitoes who survive incubation period
rho = FeverPf*TreatPf # probability of clearing infection through treatment cascade
```


We need to perform some arithmetic to convert the travel data into a TaR matrix.
```{r TaR matrix}
# get the region - to - pixel matrix
source(here("scripts/region_to_areaId_mapping.R"))
# convert the trip.dest.data into a matrix
trip.dest.mat <- as.matrix(trip.dest.data[year == 2018 & draw == "draw.mean", .(t_eg, ti_ban, ti_lub, ti_mal, ti_mok, ti_ria, ti_ure)])
trip.dest.mat <- rbind(trip.dest.mat, matrix(c(1,0,0,0,0,0,0), ncol = 7))
# take the matrix product of trip.dest.mat with the region-to-pixel matrix toget pixel-to-pixel:
movement.matrix <- trip.dest.mat %*% reg.2.pixel

# NB: the movement matrix is the probability of going from each areaId to each other areaId;
# we need to combine it with the other travel data to obtain the full TaR matrix
TaR.matrix <- diag(1, nrow = 242, ncol = 242)
# vector of trip durations across the island and off-island
trip.duration <- c(rep(trip.duration.bi, 241), trip.duration.eg)
# vector of frequencies at which people leave home
# where we divide by 56 to transform the probability of leaving into 
# the frequency of leaving during the study period
trip.freq <- c((trip.freq.data[year == 2018]$draw.mean)/56, 1)
for (i in 1:242){
  TaR.matrix[i,] <- (movement.matrix[i,]*trip.duration)/(sum(movement.matrix[i,]*trip.duration) + 1/trip.freq[i])
  TaR.matrix[i,i] <- 1 - sum(TaR.matrix[i,])
  TaR.matrix[i,] <- TaR.matrix[i,]/sum(TaR.matrix[i,])
}
#rowSums(TaR.matrix)
#diag(TaR.matrix)
## Observation: most people spent > 90% of their time at home!
```

We need to perform some arithmetic to derive the EIR based off of the TaR matrix calculated above.  At the same time we can derive lambda and sporozoite rate, which we will also need for parameterizing the simulation
```{r Derive EIR}
pfpr.input <- c(pfpr.data$draw.mean, 0.43)

odds.vector <- r/(1-rho)*pfpr.input/(1-(1+rho*r/eta/(1-rho))*pfpr.input)
h.FOI <- MASS::ginv(TaR.matrix) %*% odds.vector
h.FOI[which(h.FOI < 0)] <- 0
```


```{r Derive Initial Conditions for Mosquitoes}
# we begin by calculating the fraction of infected mosquitoes, the sporozoite rate, from the Ross-Macdonald equations + Kappa
# Total population, including visitors
H.visitors <- t(c(pop.data[year == 2018]$pop, 0) %*% TaR.matrix)
# Sick population, including visitors
X.visitors <- t((c(pop.data[year == 2018]$pop, 0) * pfpr.input)  %*% TaR.matrix)
# This is the number of people who are sick, including visitors and residents both
kappa <- X.visitors/H.visitors
kappa[242] <- .43
z.spz <- peip*a*c*kappa/(p*a*c*kappa + (1-p))
# this is Z/M, but we currently do not know M
M = h.FOI*H.visitors/a/b/z.spz

Z = z.spz * M
Z[242] = 0 # for off-island
Y = Z/peip
Y[242] = 0 # for off-island
```

```{r Derive Lambda}
# We use this quantity to set the emergence rate of mosquitoes in each of the patches
# this is Lambda, calculated based on equilibrium value for the emergence process
Lambda = M*(1-p)/p
Lambda[242] = 0 # for off-island
```



## Parameterizing:

```{r PfSI Parameters}
# vector of parameters 
pfsi_pars <- pfsi_parameters(FeverPf = 0.1116336, TreatPf = 0.602)
```

There are 242 patches in the Bioko Island model.  241 of them are the populated areaId's from the 2018 census.  The last one (indexed by 242) is a *reservoir patch* representing off-island travel on Mainland EG.
```{r Set up Patch Parameters}
# set up patches (n is how many patches we have)
n.patch <- 242 # 241 + 1 : the last patch is off-island

# set the EIR off-island
eg.eir <- h.FOI[242]/b

patch_pars <- patches_parameters(move = movement.matrix,
                                 bWeightZoo = rep(0,n.patch),
                                 bWeightZootox = rep(0,n.patch),
                                 reservoir = c(rep(F,(n.patch-1)), T),
                                 res_EIR = c(rep(0,(n.patch-1)),eg.eir)
)
```

```{r Set up Mosquito Parameters}
psi <- Matrix::sparseMatrix(i = {},j = {},x = 0.0,dims = c(n.patch,n.patch))
diag(psi) <- rep(1,n.patch)
lambda.matrix = t(matrix(Lambda, nrow = n.patch, ncol = 365))
mosy_pars <- mosquito_rm_conpars(N = n.patch,
                                 lambda = lambda.matrix,
                                 psi = psi,
                                 EIP = rep(11,365),
                                 M = M,
                                 Y = Y,
                                 Z = Z)
```


```{r Set up Human Parameters}
# human populations
# number of people in each patch
patch.human.pop <- c(pop.data[year == 2018]$pop, 0) 
# malaria prevalence in each patch
pfpr <- pfpr.input
# total number of humans
n.humans <- sum(patch.human.pop) 

# sample S or I for each person
# set seed first, for the sake of reproducibility
set.seed(42L)
init_state <- unlist(
  mapply(FUN = function(n,pr){
    sample(x = c("I","S"),size = n,replace = T,prob = c(pr,1-pr))
    },
  n=patch.human.pop,
  pr=pfpr.input,
  SIMPLIFY = F)
  )

# Define Patch IDs, for where people go
# These IDs correspond to indices in a vector - need it to be 0-indexed for c++
patch_id <- rep(0:(n.patch-1), times=patch.human.pop)

# Set uniform biting weights; this could follow any density on the positive reals (gamma, log-normal, weibull, etc.)
bweights <- rep(1,n.humans)

# Set mean trip durations based on destination
trip.durations <- c(rep(trip.duration.bi, 241), trip.duration.eg)
# Set trip frequencies for each human - this is set according to one's home origin patch
trip.freqs <- rep(trip.freq.data[year == 2018]$draw.mean/56, times = patch.human.pop[1:241])


# the data structure that will be passed down to C++ to construct the human pop
human_pars <- vector("list",n.humans)
for(i in 1:n.humans){
  human_pars[[i]] <- human_pfsi_conpars(id = i-1,
                                        home_patch_id = patch_id[i],
                                        trip_duration = trip.durations,
                                        trip_frequency = trip.freqs[i],
                                        bweight = bweights[i],
                                        age = 20,
                                        state = init_state[i],
                                        bite_algorithm = 0)
}
```


```{r Set output filenames and paths}
output_dir = here("data/sim_outputs/bioko_island_run")
if (!dir.exists(output_dir)) {dir.create(output_dir)}


log_pars <- list()
#h_inf <- paste0(path,"pfsi.csv")
h_inf <- here("data/sim_outputs/bioko_island_run/pfsi.csv")
log_pars[[1]] <- list(outfile = h_inf, 
                      key = "pfsi",
                      header = paste0(c("time",
                                        "patch",
                                        unlist(lapply(c("S","I","P"),function(x){
                                          paste0(x,c("_visitor","_resident_home","_resident_away"))}
                                          )),
                                        "incidence_resident",
                                        "incidence_traveller"),
                                      collapse = ",")
                    )
#mosy <- paste0(path,"mosy.csv")
mosy <-  here("data/sim_outputs/bioko_island_run/mosy.csv")
log_pars[[2]] <- list(outfile = mosy,
                      key = "mosquito",
                      header = paste0(c("time",
                                        "state",
                                        paste0("patch",1:n.patch)),
                                      collapse = ","))
```


```{r Run a simulation}
# Set random seed, for reproducibility
set.seed(1L)
# 
run_macro(tmax = 100,
          human_pars = human_pars,
          mosquito_pars = mosy_pars,
          patch_pars = patch_pars,
          model_pars = pfsi_pars,
          log_streams = log_pars,
          vaxx_events = NULL,
          verbose = T)
```


```{r Handling a single set of outputs}
dt <- fread(h_inf)

# Where are the patch indexes for a handful of patches?
# which(areaId.list == 220) # in Malabo
# which(areaId.list == 335) # in Malabo
# which(areaId.list == 502) # Basupu
# which(areaId.list == 1175) # in Baney
# which(areaId.list == 2199) # in Luba
# which(areaId.list == 2457) # in Riaba
# patch.human.pop[c(9, 31, 67, 121, 171, 175)]

# Create a new data table to merge onto the simulation output, to track population denominators over time
areaId.list <- sort(pop.data[year == 2018]$areaId)
pop.dt <- data.table(patch = c(0:(242-1)), areaId = areaId.list)
pop.dt <- merge(pop.dt, pop.data[year == 2018, .(areaId, pop)], by = "areaId")
# merge, to use the pop column as a denominator when calculating fractions susceptible, infected, protected
dt <- merge(dt, pop.dt, by = "patch")
dt[, s := (S_resident_home + S_resident_away)/pop, by = c("time" , "patch" , "time")]
dt[, i := (I_resident_home + I_resident_away)/pop, by = c("time" , "patch" , "time")]
dt[, p := (P_resident_home + P_resident_away)/pop, by = c("time" , "patch" , "time")]

h <- melt(dt[patch %in% (c(9, 31, 67, 121, 171, 175) - 1)],
     id.vars = c("time", "areaId"), 
     measure.vars = c("s","i","p"),
     value.name = "fraction")

ggplot(data = h) + 
  geom_point(mapping = aes(x = time, y = fraction, color = variable), shape = 20, size = .01) + 
  facet_wrap(~areaId)
```


```{r Ensemble Run}
# run ensemble
nrun <- 10
tsteps <- 100
pb <- txtProgressBar(min = 1,max = nrun)

for(k in 1:nrun){
  
  log_pars <- list()
  h_inf <- h_inf <- here("data/sim_outputs/bioko_island_run",
                         paste0("pfsi_ens_",k,".csv"))
  log_pars[[1]] <- list(outfile = h_inf, 
                        key = "pfsi",
                        header = paste0(c("time",
                                          "patch",
                                          unlist(lapply(c("S","I","P"),function(x){
                                            paste0(x,c("_visitor","_resident_home","_resident_away"))}
                                            )),
                                          "incidence_resident",
                                          "incidence_traveller"),
                                        collapse = ",")
                      )

  mosy <-  here("data/sim_outputs/bioko_island_run",
                         paste0("mosy_ens_",k,".csv"))
  log_pars[[2]] <- list(outfile = mosy,
                        key = "mosquito",
                        header = paste0(c("time",
                                          "state",
                                          paste0("patch",1:n.patch)),
                                        collapse = ","))
  
  # set seed for reproducibility
  set.seed(k)
  
  # set initial conditions
  init_state <- unlist(
    mapply(FUN = function(n,pr){
      sample(x = c("I","S"),size = n,replace = T,prob = c(pr,1-pr))
      },
    n=patch.human.pop,
    pr=pfpr.input,
    SIMPLIFY = F)
    )
  
  # the data structure that will be passed down to C++ to construct the human pop
  human_pars <- vector("list",n.humans)
  for(i in 1:n.humans){
    human_pars[[i]] <- human_pfsi_conpars(id = i-1,
                                          home_patch_id = patch_id[i],
                                          trip_duration = trip.durations,
                                          trip_frequency = trip.freqs[i],
                                          bweight = bweights[i],
                                          age = 20,
                                          state = init_state[i],
                                          bite_algorithm = 0)
  }
  
  run_macro(tmax = tsteps,
            human_pars = human_pars,
            mosquito_pars = mosy_pars,
            patch_pars = patch_pars,
            model_pars = pfsi_pars,
            log_streams = log_pars,
            vaxx_events = NULL,
            verbose = FALSE)

  setTxtProgressBar(pb,k)
}
```



```{r Analyze the Ensemble}
# Set ensemble of files to be analyzed
ensemble.file.list <- list.files(path = here("data/sim_outputs/bioko_island_run"),
                                 pattern = "pfsi_ens_[[:digit:]]+.csv")

# load in the first one
df_curr <- fread(here("data/sim_outputs/bioko_island_run",ensemble.file.list[1]))
df_curr <- merge(pop.dt, df_curr, by = "patch")
df_curr[, s := (S_resident_home + S_resident_away)/pop, by = c("time" , "patch" , "time")]
df_curr[, i := (I_resident_home + I_resident_away)/pop, by = c("time" , "patch" , "time")]
df_curr[, p := (P_resident_home + P_resident_away)/pop, by = c("time" , "patch" , "time")]
# copy it to create the corresponding matrix:
mat_curr <- as.matrix(df_curr)
# then we use that matrix to create holders to catch data for the means and standard deviations:
mat_mean <- mat_curr
mat_sd <- mat_mean
mat_sd[, 5:ncol(mat_curr)] <- mat_mean[, 5:ncol(mat_curr)]^2
# now we loop over the other files in the list of ensemble outputs
nrun = length(ensemble.file.list)
for (i in 2:nrun){
  df_curr <- fread(here("data/sim_outputs/bioko_island_run",ensemble.file.list[i]))
  
  df_curr <- merge(pop.dt, df_curr, by = "patch")
  df_curr[, s := (S_resident_home + S_resident_away)/pop, by = c("time" , "patch" , "time")]
  df_curr[, i := (I_resident_home + I_resident_away)/pop, by = c("time" , "patch" , "time")]
  df_curr[, p := (P_resident_home + P_resident_away)/pop, by = c("time" , "patch" , "time")]
  
  mat_curr <- as.matrix(df_curr)
  mat_mean[, 5:ncol(mat_curr)] <- mat_mean[, 5:ncol(mat_curr)] + mat_curr[, 5:ncol(mat_curr)]
  mat_sd[, 5:ncol(mat_curr)] <-  mat_sd[, 5:ncol(mat_curr)] + mat_curr[, 5:ncol(mat_curr)]^2
}
mat_mean <- as.data.table(mat_mean)
mat_sd <- as.data.table(mat_sd)

mat_mean[, 5:ncol(mat_curr)] <- mat_mean[, 5:ncol(mat_curr)]/nrun
mat_sd[, 5:ncol(mat_curr)] <- mat_sd[, 5:ncol(mat_curr)]/nrun - mat_mean[, 5:ncol(mat_curr)]^2
mat_sd[mat_sd < 0] <- 0
mat_sd[, 5:ncol(mat_curr)] <- sqrt(mat_sd[, 5:ncol(mat_curr)])

h <- melt(mat_mean[areaId %in% c(220,335,502,1175,2199,2457)],
     id.vars = c("time", "areaId"), 
     measure.vars = c("s","i","p"),
     value.name = "fraction")
h.sd <- melt(mat_sd[areaId %in% c(220,335,502,1175,2199,2457)],
     id.vars = c("time", "areaId"), 
     measure.vars = c("s","i","p"),
     value.name = "fraction.sd")
h <- merge(h, h.sd, by = c("time", "areaId", "variable"))

ggplot(data = h) + 
  geom_errorbar(mapping = aes(x = time, ymin = fraction - fraction.sd, ymax = fraction + fraction.sd, color = variable, alpha = 0.1)) + 
  geom_point(mapping = aes(x = time, y = fraction, color = variable), shape = 20, size = .01) + 
  facet_wrap(~areaId)
```

Numerical comparison of simulation outputs to mean pfpr's used for calibration
```{r Numerical comparison}

data.table(h[variable == "i", .(mean(fraction), mean(fraction.sd)), by = c("areaId", "variable")], input.pfpr = pfpr.input[c(9, 31, 67, 121, 171, 175)])

```

