---
title: "macro-pfsi-download"
author: "Daniel T. Citron"
date: "1/13/2020"
output: html_document
---

Run this script first - we will use this script to download the libraries that we'll need.

```{r}
library(devtools)
library(here, lib.loc = "/ihme/malaria_modeling/libraries")
library(withr)
```

We are going to use `here` to bundle everthing that you'll need together in one directory. (Note that `here` is installed in the `/malaria_modeling/libraries` location, a new directory which we will use to share libraries!)

Check your path location - I recommend setting it to the `/modeling-tutorial` directory -
```{r}
#setwd("pick/your/favorite/location")
here()
```

Install and load `macro.pfsi` from repo - default target directory will be `/libraries/macro.pfsi`.
```{r Load macro.pfsi from repo, eval=FALSE}
# Check that the target directory there:
output_dir = here("libraries")
if (!dir.exists(output_dir)) {dir.create(output_dir)}
output_dir = here("libraries/macro.pfsi")
if (!dir.exists(output_dir)) {dir.create(output_dir)}

# Download the lib
withr::with_libpaths(
  new = output_dir,
  devtools::install_github(repo = "https://github.com/dd-harp/MASH",subdir = "macro.pfsi", force = TRUE)
)

# Make sure it is possible to load the libraries
library(RcppArmadillo, lib.loc = here("libraries/macro.pfsi"))
library(macro.pfsi, lib.loc = here("libraries/macro.pfsi"))
```