---
title: "macro-pfsi-download"
author: "Daniel T. Citron"
date: "1/13/2020"
output: html_document
---

Run this script first - we will use this script to download the libraries that we'll need.

```{r}
library(devtools)
library(here)
library(withr)
```

We are going to use `here` to bundle everthing that you'll need together in one directory.

Check the current path location - if you're working inside the `MASH-group-tutorial.Rproj` then the `here` library should default to setting the current path location accordingly. If you want to change that, feel free to do so.
```{r}
#setwd("pick/your/favorite/location")
here()
```

Install and load `macro.pfsi` from repo - default target directory will be `/libraries/macro.pfsi`. Alternatively, if you are working remotely on the cluster in the `/ihme/malaria_modeling` directory you can find where I've installed it at `/ihme/malaria_modeling/dtcitron/Rlibs/macro.pfsi`.

```{r Load macro.pfsi from repo, eval=FALSE}
# Check that the target directory there:
output_dir = here("libraries/macro.pfsi")
if (!dir.exists(output_dir)) {dir.create(output_dir)}

# Download the lib
withr::with_libpaths(
  new = output_dir,
  devtools::install_github(repo = "https://github.com/dd-harp/MASH",subdir = "macro.pfsi", force = TRUE)
)

# Make sure it is possible to load the libraries
library(RcppArmadillo, lib.loc = here("libraries/macro.pfsi"))
library(macro.pfsi, lib.loc = here("libraries/macro.pfsi"))
```