---
title: "MASH-macro-workflow"
author: "Daniel T. Citron"
date: "1/13/2020"
output: html_document
---

# MACRO TUTORIAL

The purpose of this notebook will be to demonstrate how to use the `macro.pfsi` library.  We will demonstrate a workflow:

* Set up simulation: Enter model parameters which make the simulation specific
* Run a single simulation; Analyze output
* Run an ensemble of simulations; Analyze output

Things to discuss:

* Which parameters we can consider fixed; which parameters require external data to calibrate
* How to interpret simulation outputs


## Step 1: Load libraries
First run `macro-pfsi-download.Rmd` and make sure that `macro.pfsi` library is available. Make sure that you are pointing to the correct directory; if this throws errors then we will need to fix things before we proceed.
```{r Call Libraries}
library(data.table)
library(ggplot2)
library(Matrix)
# in case you need to re-load the here library:
#library(here, lib.loc = "/ihme/malaria_modeling/dtcitron/Rlibs/")

library(RcppArmadillo, lib.loc = here("libraries/macro.pfsi"))
library(macro.pfsi, lib.loc = here("libraries/macro.pfsi"))
```

Set the random seed for setting initial conditions
```{r Set random seed}
# seed
set.seed(42L)
```

## The Parameters

In the following code blocks, we will start to build the software object that lets us run a specific simulation. This will require specifying parameters. Different parameters are entered into the simulation in different places.  Right now each of those places roughly maps onto a different module:

### Course of infection:
(This can be parameterized differently in a lot of ways)
```{r Course of Infection Setup}
# vector of parameters describing course of infection - PfSI
pfsi_pars <- pfsi_parameters(mnChemoprophylaxisPf = 32, # duration of chemoprophylaxis
                             PEProtectPf = 1, # % of people the PE vaccine works in
                             peBlockPf = 1, # % decrease in transmission efficiency of new infections
                             mnPEPf = 200, vrPEPf = 50 # duration of  PE vaccine
                             )
# check on all of the other parameters we can change
pfsi_parameters()
```

### Patch setup
```{r Patch set-up}

# set up patches (n is how many patches we have)
n.patch <- 5
# movement matrix for humans is uniform over possible destinations.
move <- matrix(data = 1/(n.patch-1),nrow = n.patch, ncol = n.patch)
diag(move) <- 0
patch_pars <- patches_parameters(move = move,
                                 bWeightZoo = rep(0,n.patch),
                                 bWeightZootox = rep(0,n.patch),
                                 reservoir = rep(F,n.patch),
                                 res_EIR = rep(0,n.patch))
```

### Mosquito population setup
```{r Mosquito population setup}
## mosquito parameters
# psi is the movement matrix for mosquitoes, if we let mosquitoes drift from one place to another
psi <- Matrix::sparseMatrix(i = {},j = {},x = 0.0,dims = c(n.patch,n.patch))
diag(psi) <- rep(1,n.patch)

# lambda is the per-day rate that new adult mosquitoes emerge at each patch
# In our model, the abundance of mosquitoes drives the transmission intensity, which in turn drives PR
# (In a more complicated course-of-infection model, the relationship between mosquito abundance, incidence, and PR will be much more complicated)
# The matrix has 1 row for each value of lambda, and 1 column for each day of the year
# This allows us to specify the annual cycles in mosquito emergence rates (but does not yet account for multi-year trends)
# In this example we will have 5 patches, each of which will have a different (increasing) value of lambda
lambda.matrix = t(matrix(c(10,20,30,40,50), nrow = n.patch, ncol = 365))

# Set up mosquito parameters
mosy_pars <- mosquito_rm_conpars(N = n.patch,
                                 lambda = lambda.matrix,
                                 psi = psi,
                                 EIP = rep(11,365),
                                 M = rep(450,n.patch),
                                 Y = rep(0,n.patch),
                                 Z = rep(0,n.patch))
```

### Set up human populations
```{r Human population setup}
# humans
patch_sizes <- rep(500,n.patch) # number of people in each patch
pfpr <- rep(0.5,n.patch) # initial malaria prevalence in each patch
n.humans <- sum(patch_sizes) # total number of humans

# sample S or I for each person
# Given that each patch has 50% infected people at the start of the simulation
# we randomly assign each person in each patch to have initial state of infected or susceptible
init_state <- unlist(mapply(FUN = function(n.patch,pr){
  sample(x = c("I","S"), 
         size = n.patch,
         replace = T,
         prob = c(pr,1-pr))
  },
  n=patch_sizes,pr=pfpr,SIMPLIFY = F))

# Assign a patch ID to each human (0-indexed, because these variables go into C++)
patch_id <- rep(0:(n.patch-1),times=patch_sizes)

# Assign biting weights to humans
# uniform biting weights; this could follow any density on the positive reals (gamma, log-normal, weibull, etc.)
bweights <- rep(1,n.humans)

# Human parameters - Define parameters to each human
human_pars <- vector("list",n.humans)
for(i in 1:n.humans){
  human_pars[[i]] <- human_pfsi_conpars(id = i-1,
                                        home_patch_id = patch_id[i],
                                        # assign duration of travel of 3 days to each destination patch to all humans
                                        trip_duration = rep(3,n.patch), 
                                        # assign rate of travel to once every 20 days to all humans
                                        trip_frequency = 1/20, 
                                        bweight = bweights[i],
                                        # relevant if the course-of-infection model had lifetime exposure dependence; not relevant for PfSI
                                        age = 20, 
                                        state = init_state[i], # initial state of the ith human
                                        bite_algorithm = 0) # dunno what this one is
}

```

### Vaccines/interventions
Set up the schedule for vaccinations, with or without accompanying treatment
```{r Vaccinations setup}
# Vaccinations - this is an example of scheduling an intervention

# vaxx_id comes up with a sample of 25 humans to treat
vaxx_id <- sample(x = 0:(n.humans-1),size = n.humans*0.25,replace = F)
# Treat everyone
vaxx_pars <- lapply(X = vaxx_id, FUN = function(id){
  vaccination_pfsi_conpars(id = id, # who to vaccinate
                           t = 500, # vaccinate on day 500
                           treat = T,
                           type = "PE")
})

# Similarly, you can set this list to be empty if you want no interventions
#vaxx_pars <- list()

```


### Define outputs
We specify the directory location for storing our output files, and then create some documents to contain the data as we output it.
```{r Set up a place to save all of the output files}
log_pars <- list()

# Where to put the human outputs
human.out.path <- here("data/sim_outputs/pfsi.csv")
log_pars[[1]] <- list(outfile = human.out.path,key = "pfsi",
                      header = paste0(c("time","patch",unlist(lapply(c("S","I","P"),function(x){paste0(x,c("_visitor","_resident_home","_resident_away"))})),"incidence_resident","incidence_traveller"),collapse = ",")
                    )

# Where to put the mosquito outputs
mosy.out.path <- here("data/sim_outputs/mosy.csv")
log_pars[[2]] <- list(outfile = mosy.out.path,key = "mosquito",
                      header = paste0(c("time","state",paste0("patch",1:n.patch)),collapse = ","))
```

## Run the simulation:

```{r Run the simulation}
run_macro(tmax = 1e3,
          human_pars = human_pars,
          mosquito_pars = mosy_pars,
          patch_pars = patch_pars,
          model_pars = pfsi_pars,
          log_streams = log_pars,
          vaxx_events = vaxx_pars,
          verbose = T)
```

# Postprocess and visualize output for analysis
Postprocessing is very tricky, and only gets more complicated and computationally intensive for very large and very complicated simulations.

Examine the human population data 
```{r}
pfsi.dat <- fread(human.out.path)
View(pfsi.dat)
```

Examine the mosquito population data
```{r}
mosy.dat <- fread(mosy.out.path)
View(mosy.dat)
```

Discuss: what does this graph mean?
```{r Handling a single set of outputs}
# List of patches (named according to their index number)
patches.list <- c(1:5)
# Population for each patch
pop.dt <- data.table(patch = patches.list, pop = rep(500,n.patch))

# Note that patches are zero-indexed, so we'll change it by one...
pfsi.dat$patch <- pfsi.dat$patch + 1
pfsi.dat <- merge(pfsi.dat, pop.dt, by = "patch")

pfsi.dat[, s := (S_resident_home + S_resident_away)/pop, by = c("time" , "patch" , "time")]
pfsi.dat[, i := (I_resident_home + I_resident_away)/pop, by = c("time" , "patch" , "time")]
pfsi.dat[, p := (P_resident_home + P_resident_away)/pop, by = c("time" , "patch" , "time")]

h <- melt(pfsi.dat,
     id.vars = c("time", "patch"), 
     measure.vars = c("s","i","p"),
     value.name = "fraction")

ggplot(data = h) + 
  geom_point(mapping = aes(x = time, y = fraction, color = variable), shape = 20, size = .01) + 
  facet_wrap(~patch)
```

(Another way to handle postprocessing, brought to you by tidyverse)
```{r Tidyverse postprocessing}
#pfsi <- readr::read_csv(h_inf)

# library(tidyverse)
# pfsi_pr <- pfsi %>%
#   select(-ends_with("away")) %>%
#   select(-starts_with("incidence")) %>%
#   gather(key, value, -time,-patch)
# 
# ggplot(pfsi_pr) +
#   geom_line(aes(x=time,y=value,color=key)) +
#   facet_wrap(. ~ patch) +
#   theme_bw()
```



## Ensemble modeling
It isn't enough to produce a single simulation output - our simulations are stochastic and the outputs are noisy. We can characterize this noise by simulating a big bunch - an ensemble - of different simulation "runs" where each run has the same parameters but a different random number seed for starting the initial conditions.

```{r Run Ensemble}
# Set number of simulation runs to 10
nrun <- 10
# Simulate for 1 year
tsteps <- 365
# Progress bar, for convenient perusal
pb <- txtProgressBar(min = 1,max = nrun)

output_dir = here("data/sim_outputs/ensemble_run")
if (!dir.exists(output_dir)) {dir.create(output_dir)}

## Loop over runs:
# Set output location
# Set random seed each time
# Set initial conditions each time
for(i in 1:nrun){
  # set output location
  log_pars <- list()
  human.out.path <- here("data/sim_outputs/ensemble_run", paste0("pfsi_ens_",i,".csv"))
  log_pars[[1]] <- list(outfile = human.out.path, 
                        key = "pfsi",
                        header = paste0(c("time",
                                          "patch",
                                          unlist(lapply(c("S","I","P"),function(x){
                                            paste0(x,c("_visitor","_resident_home","_resident_away"))}
                                            )),
                                          "incidence_resident",
                                          "incidence_traveller"),
                                        collapse = ",")
                      )

  mosy.out.path <-  here("data/sim_outputs/ensemble_run", paste0("mosy_ens_",i,".csv"))
  log_pars[[2]] <- list(outfile = mosy.out.path,
                        key = "mosquito",
                        header = paste0(c("time",
                                          "state",
                                          paste0("patch",1:n.patch)),
                                        collapse = ","))
  
  # set seed for reproducibility
  set.seed(i)
  
  # set initial conditions
  init_state <- unlist(mapply(FUN = function(n.patch,pr){
  sample(x = c("I","S"), 
         size = n.patch,
         replace = T,
         prob = c(pr,1-pr))
    },
    n=patch_sizes, 
    pr=pfpr, 
    SIMPLIFY = F)
  )

  # the data structure that will be passed down to C++ to construct the human pop
  human_pars <- vector("list",n.humans)
  for(j in 1:n.humans){
    human_pars[[j]] <- human_pfsi_conpars(id = j-1,
                                          home_patch_id = patch_id[j],
                                          trip_duration = rep(3,n.patch), 
                                          trip_frequency = 1/20, 
                                          bweight = bweights[j],
                                          age = 20, 
                                          state = init_state[j],
                                          bite_algorithm = 0)
  }
  
  run_macro(tmax = tsteps,
            human_pars = human_pars,
            mosquito_pars = mosy_pars,
            patch_pars = patch_pars,
            model_pars = pfsi_pars,
            log_streams = log_pars,
            vaxx_events = NULL,
            verbose = FALSE)
  
    setTxtProgressBar(pb,i)
}
```

Instead of analyzing a single simulation runs, we now have to examine an ensemble of 10 simulation runs. We can calculate statistics across the ensemble - mean and standard deviation, for example.
```{r}
ensemble.file.list <- list.files(path = here("data/sim_outputs/ensemble_run"), 
                                 pattern = "pfsi_ens_[[:digit:]]+.csv")
# load in the first one
df_curr <- fread(here("data/sim_outputs/ensemble_run",ensemble.file.list[1]))
# copy it to create the corresponding matrix:
mat_curr <- as.matrix(df_curr)
# then we use that matrix to create holders to catch data for the means and standard deviations:
mat_mean <- mat_curr[, 3:ncol(mat_curr)]
mat_sd <-  mat_mean^2

# now we loop over the other files in the list of ensemble outputs
nrun = length(ensemble.file.list)
for (i in 2:nrun){
  df_curr <-  fread(here("data/sim_outputs/ensemble_run",ensemble.file.list[i]))
  mat_curr <- as.matrix(df_curr)
  mat_mean <- mat_mean + mat_curr[, 3:ncol(mat_curr)]
  mat_sd <-  mat_sd + mat_curr[, 3:ncol(mat_curr)]^2
}

# A kludge for how we calculate mean and standard deviation, without needing to load in 
# all outputs at the same time:
mat_mean <- mat_mean/nrun
mat_sd <- sqrt(mat_sd/nrun - mat_mean^2)

mat_mean <- cbind(df_curr[,1:2], as.data.table(mat_mean))
mat_sd <- cbind(df_curr[,1:2], as.data.table(mat_sd))
```

Plot, with error bars:
```{r}
h <- melt(mat_mean, 
     id.vars = c("time", "patch"), 
     measure.vars = c("S_visitor","S_resident_home","S_resident_away",
                      "I_visitor","I_resident_home","I_resident_away",
                      "P_visitor","P_resident_home","P_resident_away"),
     value.name = "counts")

h.sd <- melt(mat_sd, 
     id.vars = c("time", "patch"), 
     measure.vars = c("S_visitor","S_resident_home","S_resident_away",
                      "I_visitor","I_resident_home","I_resident_away",
                      "P_visitor","P_resident_home","P_resident_away"),
     value.name = "counts.sd")

h <- merge(h, h.sd, by = c("time", "patch", "variable"))

ggplot(h) + 
  geom_errorbar(mapping = aes(x = time, ymin = counts - counts.sd, ymax = counts + counts.sd, color = variable, alpha = .1)) +
  geom_point(mapping = aes(x = time, y = counts), size = .01) + 
  facet_wrap(~patch)
```


Another way of analyzing the ensemble of simulation runs, brought you by Tidyverse:
```{r Tidyverse (old) post-processing}
# library(tidyverse)
# 
# pfsi_ensemble <-
#   list.files(path = here("data/sim_outputs/ensemble_run"),pattern = "pfsi_ens_[[:digit:]]+.csv") %>%
#   map_df(~read_csv(paste0(here("data/sim_outputs/ensemble_run/"),.)),.id = "run")
# 
# pfsi_ensemble_pr <- pfsi_ensemble %>%
#   select(-ends_with("away")) %>%
#   select(-starts_with("incidence")) %>%
#   gather(key, value, -time,-patch,-run)
# 
# ggplot(pfsi_ensemble_pr,aes(x=time,y=value,color=key,fill=key)) +
#   stat_summary(fun.data = median_hilow,fun.args = list(conf.int = 0.95),geom = "ribbon",alpha=0.4,color=NA) +
#   stat_summary(geom="line", fun.y="mean") +
#   facet_wrap(. ~ patch) +
#   guides(color = FALSE) +
#   theme_bw()
```